---
title: "E2 VoC Analyses Part 2: Regression Analyses"
date: 3/27/24
output:
    html_document:
        df_print: 'paged'
        toc: true
        toc_float:
            collapsed: false
            smooth_scroll: true
        number_sections: false
        code_download: true
        self_contained: true
---

```{r chunk settings, include = FALSE}
# set chunk settings
knitr::opts_chunk$set(echo = FALSE, 
                      cache = TRUE,
                      message = FALSE,
                      warning = FALSE)
knitr::opts_chunk$set(dpi=600)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```

```{r load libraries, include = F}

#load libraries
library(tidyverse)
library(glue)
library(afex)
library(sjPlot)

#load scripts
source('analysis_scripts/voc_functions.R')
```

```{r import data}

# read in learning data
learning_data <- read_csv('data/processed/learning_data.csv')

# read in participant ages
participant_ages <- read_csv('data/voc_sub_info.csv') 

# join
learning_data <- inner_join(learning_data, participant_ages, by = c('subject_id')) %>%
  mutate(age_group = case_when(age < 13 ~ 'Children',
                               age < 18 & age > 12.99 ~ 'Adolescents',
                               age > 18 ~ 'Adults'))

learning_data$age_group <- factor(learning_data$age_group,
                                  levels = c("Children", "Adolescents", "Adults"))

```

```{r process learning data}
learning_data <- learning_data %>%
  mutate(ev_choice = case_when(context == 0 ~ 9,
                               context == 1 ~ 7,
                               context == 2 ~ 5),
         ev_comp = 5 + offer,
         voc = ev_choice - ev_comp,
         better_machine = case_when(reward_prob_L > reward_prob_R ~ 1,
                                    reward_prob_L < reward_prob_R ~ 0,
         ),
         stage_2_acc = case_when(stage_2_choice == better_machine ~ 1,
                                 stage_2_choice != better_machine ~ 0)) %>%
  group_by(subject_id, context) %>%
  mutate(condition_trial = rank(trial),
         block = floor((trial-1)/21 + 1))

# exclude first-stage misses and first-stage RT < 150 ms
learning_data_filtered <- learning_data %>%
  filter(stage_1_rt > 150)

```

# Participant info
```{r subject information}
sub_info <- learning_data_filtered %>%
  ungroup() %>%
  select(subject_id, age, age_group, gender) %>%
  unique() %>%
  group_by(age_group) %>%
  summarize(N = n(), 
            min_age = min(age, na.rm = T),
            max_age = max(age, na.rm = T),
            mean_age = mean(age, na.rm = T),
            sd_age = sd(age, na.rm = T),
            n_female = sum(gender == 'Female'),
            n_male = sum(gender == 'Male'),
            n_other = sum(gender == 'Other'))
sub_info

```


# Agency task: Agency decisions 
## Model: Agency decisions by VoC
```{r agency model}
# select relevant variables 
agency_model_data <- learning_data_filtered %>%
  select(subject_id, stage_1_choice, voc, condition_trial, block, trial, age, age_group)

## REGRESSION MODEL ##
#z score continuous variables
agency_model_data$subject_id <- factor(agency_model_data$subject_id)
agency_model_data$voc_z <- scale_this(agency_model_data$voc)
agency_model_data$condition_trial <- scale_this(agency_model_data$condition_trial)
agency_model_data$age_z <- scale_this(agency_model_data$age)

#run model
agency_model <- mixed(stage_1_choice ~  voc_z * condition_trial * age_z + (voc_z * condition_trial || subject_id),
                      data = agency_model_data,
                      family = "binomial",
                      method = "LRT",
                      expand_re = T,
                      control=glmerControl(optimizer="bobyqa", optCtrl=list(maxfun=1e6)))

#show model results
agency_model
summary(agency_model)
```


```{r agency model print model stats}

agency_model.glmer =  mixed(stage_1_choice ~  voc_z * condition_trial * age_z + (voc_z * condition_trial || subject_id),
                      data = agency_model_data,
                      family = "binomial",
                      method = "LRT",
                      expand_re = T,
                      control=glmerControl(optimizer="bobyqa", optCtrl=list(maxfun=1e6)),
                      return = "merMod")

tab_model(agency_model.glmer, 
          pred.labels = c("intercept", "VoC", "trial", "age", "VoC x trial", "VoC x age", "trial x age", "VoC x trial x age"),
          transform = NULL,
          show.est = T, 
          show.se = T, 
          show.stat = T,
          show.ci = F,
          show.re.var = F,
          show.icc = F,
          show.ngroups = F,
          show.obs = F,
          show.r2 = F,
          string.se = "SE",
          emph.p = F,
          string.pred = "Predictor",
          title = "",
          dv.labels = "")
```

## Plot: Sensitivity to the value of choice
```{r voc plot, fig.height = 4, fig.width = 7, unit = "in"}
## PLOT ##
agency_sub_means <- agency_model_data %>% 
  mutate(task_half = case_when(trial < 158 ~ "First Half of Task",
                              trial > 157 ~ "Second Half of Task")) %>%
  group_by(task_half, voc, subject_id, age_group) %>%
  summarize(mean_sub_agency = mean(stage_1_choice, na.rm = T))

agency_means <- agency_sub_means %>% 
  group_by(task_half, voc, age_group) %>%
  summarize(mean_agency = mean(mean_sub_agency, na.rm = T),
            se_agency = sd(mean_sub_agency / sqrt(n())))

agency_plot <- ggplot(agency_means, aes(x = voc, y = mean_agency, color = age_group)) +
  facet_wrap(~task_half) +
  geom_point(aes(color = age_group)) + 
  geom_errorbar(aes(ymin = mean_agency - se_agency, ymax = mean_agency + se_agency), width = .1) + 
  geom_line() +
  voc_theme() + 
    scale_color_manual(values=c("#702963", "#c00000", "#ffae42"), name = "Age Group") +
  xlab("Value of Choice (VoC)") +
  ylab("Proportion Agency Choices") +
  geom_hline(yintercept = .5, linetype = "dashed") +
  geom_vline(xintercept = 0, linetype = "dashed")
agency_plot
```


## Plot: Sensitivity to value of choice with continuous age 
```{r voc plot continuous age, fig.height = 3.9, fig.width = 3, unit = "in"}

#run model without age to get random effects for each participant
agency_glmer <- mixed(stage_1_choice ~  voc_z * condition_trial + (voc_z * condition_trial | subject_id),
                      data = agency_model_data, 
                      family = binomial, 
                      method = "LRT",
                      control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=1e6)),
                      return = "merMod") 

#get fixed effect of zVoC
VoC_fixedeff <- as.data.frame(coef(summary(agency_glmer)))$Estimate[2]
VoC_int_fixedeff <- as.data.frame(coef(summary(agency_glmer)))$Estimate[4]

#get random effects
VoC_effects <- ranef(agency_glmer)$subject_id %>%
    rownames_to_column(var = "subject_id")

#combine with age
VoC_subEffects <- agency_model_data %>%
    select(subject_id, age) %>% 
    unique() %>%
    left_join(VoC_effects, by = c("subject_id")) %>%
    mutate(zVoCFull = voc_z + VoC_fixedeff, 
           intFull = `voc_z:condition_trial` + VoC_int_fixedeff)

#plot age by VoC effect
VoC_plot_continuousAge <- ggplot(VoC_subEffects, aes(x = age, y = zVoCFull)) +
    geom_point(color = "black") + 
    geom_smooth(method = "lm", color = "black", fill = "black") +
    voc_theme() + 
    xlab("Age") +
    ylab("VoC Effect") 
VoC_plot_continuousAge

#plot age by VoC x trial effect
VoC_plot_continuousAgeTrial <- ggplot(VoC_subEffects, aes(x = age, y = intFull)) +
    geom_point(color = "black") + 
    geom_smooth(method = "lm", color = "black", fill = "black") +
    voc_theme() + 
    xlab("Age") +
    ylab("VoC x Trial Effect") 
VoC_plot_continuousAgeTrial
```



## Summary stats: Sensitivity to value of control
```{r voc summary stats}

# What proportion of trials did participants choose agency when VoC was 0?
VoC_zero_means_sub <- learning_data_filtered %>% 
    filter(voc == 0) %>%
    group_by(subject_id, age_group) %>%
    summarize(meanSubAgency = mean(stage_1_choice, na.rm = T))

VoC_zero_means <- VoC_zero_means_sub %>%
  ungroup() %>%
  summarize(meanAgency = mean(meanSubAgency, na.rm = T),
              seAgency = sd(meanSubAgency / sqrt(n())))
VoC_zero_means
```


## Agency decision reaction times
```{r agency RT data processing}

#compute RT
learning_data_filtered$agencyRT <- learning_data_filtered$stage_1_rt / 1000

#how many RTs faster than 100 ms?
fastRTs <- learning_data_filtered %>%
    filter(agencyRT < .1) %>%
    nrow()

#0 excluded

slowRTs <- learning_data_filtered %>%
    filter(agencyRT > 30) %>%
    nrow()

#0 excluded

banditTask.RT.data <- learning_data_filtered %>%
    filter(agencyRT > .1) %>%
    filter(agencyRT < 30) %>%
    mutate(agencyLogRT = log(agencyRT))
```

## Model: Agency RTs by VoC
```{r voc RT model}

#compute abs(voc) variable
banditTask.RT.data$absVoC <- abs(banditTask.RT.data$voc)

#scale variables
banditTask.RT.data$zAge <- scale(banditTask.RT.data$age)
banditTask.RT.data$zVoC <- scale(banditTask.RT.data$voc)
banditTask.RT.data$zAbsVoC <- scale(banditTask.RT.data$absVoC)
banditTask.RT.data$zTrialOfCond <- scale(banditTask.RT.data$condition_trial)

agency.RT.model <- mixed(agencyLogRT ~ zAge * zAbsVoC * zTrialOfCond + (zAbsVoC * zTrialOfCond | subject_id),
                         data = banditTask.RT.data,
                         method = "S", 
                         control=lmerControl(optimizer="bobyqa", optCtrl=list(maxfun=1e6)))

agency.RT.model
summary(agency.RT.model)

#effects of
# age
# VoC
# trial of condition

#no interactions
```


```{r agency RTs model print model stats}

agency.RT.lmer <- mixed(agencyLogRT ~ zAge * zAbsVoC * zTrialOfCond + (zAbsVoC * zTrialOfCond | subject_id),
                         data = banditTask.RT.data,
                         method = "S", 
                         control=lmerControl(optimizer="bobyqa", optCtrl=list(maxfun=1e6)),
                         return = "merMod")

tab_model(agency.RT.lmer, 
          pred.labels = c("intercept", "age", "VoC magnitude", "trial", "age x VoC magnitude", "age x trial", "VoC magnitude x trial", "age x VoC magnitude x trial"),
          transform = NULL,
          show.est = T, 
          show.se = T, 
          show.fstat = T,
          show.ci = F,
          show.re.var = F,
          show.icc = F,
          show.ngroups = F,
          show.obs = F,
          show.r2 = F,
          string.se = "SE",
          emph.p = F,
          string.pred = "Predictor",
          title = "",
          dv.labels = "")
```

## Plot: Agency RTs by VoC
```{r agency RT by VoC, fig.height = 4, fig.width = 8, units = "in"}

agencyRT.means.sub <- banditTask.RT.data %>%
    group_by(voc, age_group, subject_id) %>%
    summarize(meanSubRT = mean(agencyRT),
              meanSubLogRT = mean(agencyLogRT))

#compute means and SDs
agencyRT.means <- agencyRT.means.sub  %>%
    group_by(voc, age_group) %>%
    summarize(meanRT = mean(meanSubRT),
              meanLogRT = mean(meanSubLogRT),
              seRT = sd(meanSubRT)/ sqrt(n()),
              seLogRT = sd(meanSubLogRT)/ sqrt(n()))

agencyRT.VoC.plot <- ggplot(agencyRT.means, 
                            aes(x = voc, y = meanRT)) +
    facet_wrap(~age_group) +
    geom_point(size = 3, aes(color= age_group)) +
    geom_errorbar(aes(ymin = meanRT - seRT, ymax = meanRT + seRT, color = age_group), width = .1, position = position_dodge(width = .9)) +
    scale_color_manual(values = c(color1, color2, color3)) +
    xlab("Value of Choice") +
     ylab("Mean Agency Decision Time (s)") +
    voc_theme() +
    theme(legend.position = "none")
agencyRT.VoC.plot
```

# Agency task: Machine selection
## Model: Optimal machine choices across trials by condition and age
```{r machine selection decisions}
# select variables for inclusion in mixed-effects model (no age for now)
machine_model_data <- learning_data_filtered %>%
  filter(stage_1_choice == 1) %>%
  filter(context < 2) %>%
  select(subject_id, stage_2_acc, context, condition_trial, block, age, age_group) %>%
  drop_na()

## REGRESSION MODEL ##
#z score continuous variables
machine_model_data$subject_id <- factor(machine_model_data$subject_id)
machine_model_data$context <- factor(machine_model_data$context)
machine_model_data$condition_trial <- scale_this(machine_model_data$condition_trial)
machine_model_data$age_z <- scale_this(machine_model_data$age)

#run model
machine_model <- mixed(stage_2_acc ~  context * condition_trial * age_z + (context * condition_trial || subject_id),
                      data = machine_model_data,
                      family = "binomial",
                      method = "LRT",
                      expand_re = T,
                      control=glmerControl(optimizer="bobyqa", optCtrl=list(maxfun=1e6)))

#show model results
machine_model
summary(machine_model)
```


```{r machine model print model stats}

machine_model.glmer <- mixed(stage_2_acc ~ context * condition_trial * age_z + (context * condition_trial || subject_id),
                      data = machine_model_data,
                      family = "binomial",
                      method = "LRT",
                      expand_re = T,
                      control=glmerControl(optimizer="bobyqa", optCtrl=list(maxfun=1e6)),
                      return = "merMod")

tab_model(machine_model.glmer, 
          pred.labels = c("intercept", "condition", "trial", "age", "condition x trial", "condition x age", "trial x age", "condition x trial x age"),
          transform = NULL,
          show.est = T, 
          show.se = T, 
          show.stat = T,
          show.ci = F,
          show.re.var = F,
          show.icc = F,
          show.ngroups = F,
          show.obs = F,
          show.r2 = F,
          string.se = "SE",
          emph.p = F,
          string.pred = "Predictor",
          title = "",
          dv.labels = "")
```

## Plot: Proportion optimal machine selections across age groups and trials
```{r plot bandit choices across trials, width = 7, height = 4, unit = "in"}

## PLOT ##
machine_selection_sub_means <- machine_model_data %>%
  group_by(context, block, subject_id, age_group) %>% 
  summarize(sub_acc = mean(stage_2_acc, na.rm = T))

machine_selection_means <- machine_selection_sub_means %>%
  group_by(context, block, age_group) %>% 
  summarize(mean_acc = mean(sub_acc),
            se = sd(sub_acc)/sqrt(n()))

machine_selection_plot <- ggplot(machine_selection_means, aes(x=block, y=mean_acc, color=factor(context))) +
  facet_wrap(~age_group) +
  geom_point(size = 3) +
  geom_jitter(data = machine_selection_sub_means,  aes(x=block, y=sub_acc, color=factor(context)), size = .5) +
  geom_smooth(method = "lm", aes(fill = factor(context))) +
  geom_hline(yintercept = .5, linetype="dashed") +
  ylab("Proportion Optimal Machine Selections") +
  xlab("Block") +
  scale_x_continuous(breaks = c(4, 8, 12)) +
  scale_fill_manual(name="Context",
                    labels=c("90/10",
                             "70/30"),
                    values=c(color1, color3), 
                    guide = guide_legend(reverse=TRUE)) +
  scale_color_manual(name="Context",
                     labels=c("90/10",
                              "70/30"),
                     values=c(color1, color3),
                     guide = guide_legend(reverse=TRUE)) +
  voc_theme() +
  theme(strip.text = element_text(size=12))
machine_selection_plot
```




# Explicit reward knowledge task 
## Explicit reward knowledge task: summary stats
```{r explicit knowledge task}

# Read in data
explicitKnow <- read_csv('data/processed/explicit_data.csv') 
#explicitKnow$subject_id <- factor(explicitKnow$subject_id)

#combine with age
explicitKnow <- full_join(explicitKnow, participant_ages, by = c("subject_id"))

explicitKnow %>% 
  group_by(subject_id, age) %>% 
  summarize(m = mean(error)) %>% 
  ungroup() %>% 
  summarize(meanErr = mean(m, na.rm=T), sd = sd(m, na.rm = T))
```

## Model: Explicit reward knowledge by age and true probabilities
```{r explicit knowledge model}

#re-scale age and zTrueProb
explicitKnow.filtered <- explicitKnow %>%
    select(subject_id, age, true_prob, error) %>%
    drop_na()

explicitKnow.filtered$zAge <- scale(explicitKnow.filtered$age)
explicitKnow.filtered$zTrueProb <- scale(explicitKnow.filtered$true_prob)

# run model
explicitKnow_errorbyTrueProbAge.mixed <- mixed(error ~ zTrueProb*zAge + (1|subject_id), 
                                               data = explicitKnow.filtered,
                                               method = "S") 
explicitKnow_errorbyTrueProbAge.mixed
summary(explicitKnow_errorbyTrueProbAge.mixed)
```

## Plot: Explicit reward knowledge
```{r plot explicit knowledge}

explicitKnow <- explicitKnow %>%
  mutate(age_group = case_when(age < 13 ~ 'Children',
                               age < 18 & age > 12.99 ~ 'Adolescents',
                               age > 18 ~ 'Adults'))

explicitKnow$age_group <- factor(explicitKnow$age_group,
                                  levels = c("Children", "Adolescents", "Adults"))

# plot response by bandit
explicitKnow %>% drop_na() %>%
    ggplot(., aes(x=factor(true_prob), y=response, fill=age_group)) +
    geom_boxplot() +
    scale_fill_manual(values = c(color1, color2, color3), name = "Age Group") +
    ylab("Reported Reward Probability") +
    xlab("True Reward Probability") +
    scale_x_discrete(labels = c("10%", "30%", "50%", "70%", "90%")) +
    scale_y_continuous(breaks = c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10), 
                     labels = c("10%", "20%", "30%", "40%", "50%", "60%", "70%", "80%", "90%", "100%")) +
    voc_theme()
```
