---
title: "VoC Analyses Part 2: Regressions"
date: 3/27/24
output:
    html_document:
        df_print: 'paged'
        toc: true
        toc_float:
            collapsed: false
            smooth_scroll: true
        number_sections: false
        code_download: true
        self_contained: true
---
```{r chunk settings, include = FALSE}
# set chunk settings
knitr::opts_chunk$set(echo = FALSE, 
                      cache = TRUE,
                      message = FALSE,
                      warning = FALSE)
knitr::opts_chunk$set(dpi=600)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```
                      
```{r setup, include=FALSE}

# list all packages required for the analysis
list_of_packages <- c("tidyverse", "afex", "pander", "sjPlot")

# load all packages 
lapply(list_of_packages, library, character.only = TRUE)

# add theme for plotting
voc_theme <- function () {
  theme(
    panel.border = element_rect(fill = "transparent", color="gray75"),
    panel.background  = element_blank(),
    plot.background = element_blank(), 
    legend.background = element_rect(fill="transparent", colour=NA),
    legend.key = element_rect(fill="transparent", colour=NA),
    line = element_blank(),
    axis.ticks = element_line(color="gray75"),
    text=element_text(family="Avenir"),
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 15),
    title = element_text(size = 15),
    strip.background = element_blank(),
    strip.text = element_text(size=12)
  )
}

color1 = "#00b4d8"
color2 = "#0077b6"
color3 = "#03045e"


#z-score function
scale_this <- function(x){
  (x - mean(x, na.rm=TRUE)) / sd(x, na.rm=TRUE)
}

```

# Participant info
```{r participants plot}

#load demographic info
sub_info <- read_csv('data/voc_sub_info.csv') 

# plot histogram of male and female participants
sub_info %>% mutate(whole_age = floor(age)) %>% 
    group_by(subID, gender, whole_age) %>% 
    distinct(subID) %>% 
    ggplot(., aes(x=whole_age, fill=gender)) +
    geom_histogram(binwidth = 1, color="white") +
    scale_fill_manual(name="Sex",
                    labels=c("Female", "Male"),
                    values=c(color1, color2)) +
    scale_y_continuous(breaks = c(2,4,6,8,10),
                   labels = c("2","4","6","8","10"),
                   limits = c(0,10)) +
    xlab("Age") +
    ylab("Count") +
    voc_theme()
```

```{r participant info}

#load demographic info
sub_info <- read_csv('data/voc_sub_info.csv') %>%
    mutate(age_group = case_when(age < 13 ~ "Children",
                                 age > 12.99 & age < 18 ~ "Adolescents",
                                 age > 17.99 ~ "Adults"))

sub_info$age_group <- factor(sub_info$age_group, levels = c("Children", "Adolescents", "Adults"))

# age group information
age_group_info <- sub_info %>%
    group_by(age_group) %>%
    summarize(N = n(), 
              meanAge = mean(age),
              sdAge = sd(age),
              nFemale = sum(gender == "F")
              )

pander(age_group_info)
```


# Agency task: Machine selection
## Model: Optimal machine choices across trials by condition and age
```{r machine choices across trials by age}

# Read in data
banditTask <- read_csv('data/processed/bandit_task.csv') 

#combine with participant age
banditTask <- full_join(banditTask, sub_info, by = c("subID"))

# Filter data to have only trials where people choose agency and exclude trials with 50-50 condition 
banditTask.filtered <- banditTask %>% 
    filter(agency == 1, condition!="bandits5050")

# Scale continuous variables
banditTask.filtered$zAge <- scale_this(banditTask.filtered$age)
banditTask.filtered$zTrialOfCond <- scale_this(banditTask.filtered$trialOfCond)

# Mixed-effects logistic regression model
correct_byConditionTrialAge.mixed <- mixed(stage_2_acc ~ condition*zTrialOfCond*(zAge) + (condition*zTrialOfCond|subID), 
                data = banditTask.filtered,
                family = binomial, 
                method = "LRT",
                control=glmerControl(optimizer="bobyqa", optCtrl=list(maxfun=1e6)))

#display model stats
correct_byConditionTrialAge.mixed 
```

```{r machine model print model stats}

correct_byConditionTrialAge.glmer <- mixed(stage_2_acc ~ condition*zTrialOfCond*(zAge) + (condition*zTrialOfCond|subID), 
                data = banditTask.filtered,
                family = binomial, 
                method = "LRT",
                control=glmerControl(optimizer="bobyqa", optCtrl=list(maxfun=1e6)), 
                return = "merMod")

tab_model(correct_byConditionTrialAge.glmer, 
          pred.labels = c("intercept", "condition", "trial", "age", "condition x trial", "condition x age", "trial x age", "condition x trial x age"),
          transform = NULL,
          show.est = T, 
          show.se = T, 
          show.stat = F,
          show.ci = F,
          show.re.var = F,
          show.icc = F,
          show.ngroups = F,
          show.obs = F,
          show.r2 = F,
          string.se = "SE",
          emph.p = F,
          string.pred = "Predictor",
          title = "",
          dv.labels = "")
```

## Plot: Proportion optimal machine selections across age groups and trials
```{r plot bandit choices across trials, width = 7, height = 4, unit = "in"}

banditTaskSubMeans <- banditTask %>%
    mutate(block = floor((trial-1)/21) + 1) %>%
    filter(agency==1, condition!="bandits5050") %>% 
    group_by(condition, block, age_group, subID) %>% 
    summarize(pctCorrect = mean(stage_2_acc))

banditTaskMeans <- banditTaskSubMeans %>%
    group_by(condition, block, age_group) %>% 
    summarize(pctCorr = mean(pctCorrect),
              se = sd(pctCorrect)/sqrt(n()))

machineSelectionPlot <- ggplot(banditTaskMeans, aes(x=block, y=pctCorr, color=condition)) +
    facet_wrap(~age_group) +
    geom_point(size = 3) +
    geom_jitter(data = banditTaskSubMeans,  aes(x = block, y = pctCorrect, color=condition), size = .5) +
    geom_smooth(method = "lm", aes(fill = condition)) +
    geom_hline(yintercept = .5, linetype="dashed") +
    ylab("Proportion Optimal Machine Selections") +
    xlab("Block") +
    scale_x_continuous(breaks = c(4, 8, 12)) +
    scale_fill_manual(name="Condition",
                      labels=c("70/30",
                               "90/10"),
                      values=c(color1, color3), 
                      guide = guide_legend(reverse=TRUE)) +
    scale_color_manual(name="Condition",
                      labels=c("70/30",
                               "90/10"),
                      values=c(color1, color3),
                     guide = guide_legend(reverse=TRUE)) +
    voc_theme() +
    theme(strip.text = element_text(size=12))
machineSelectionPlot
```


# Agency task: Agency decisions 
## Model: Agency decisions by VoC
```{r voc model}

#scale variables
banditTask$zVoC <- scale_this(banditTask$voc)
banditTask$zTrialOfCond <- scale_this(banditTask$trialOfCond)
banditTask$zAge <- scale_this(banditTask$age)

# predict agency choice from utility of control, trial, linear age
agency_byVOCTrialAge.mixed = mixed(agency ~ zVoC * zTrialOfCond * zAge + (zVoC * zTrialOfCond|subID), 
                        data = banditTask, 
                        family = binomial, 
                        method = "LRT", control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=1e6))) 

#display stats
agency_byVOCTrialAge.mixed
```

```{r agency model print model stats}

agency_byVOCTrialAge.glmer = mixed(agency ~ zVoC * zTrialOfCond * zAge + (zVoC * zTrialOfCond|subID), 
                        data = banditTask, 
                        family = binomial, 
                        method = "LRT", control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=1e6)),
                        return = "merMod") 

tab_model(agency_byVOCTrialAge.glmer, 
          pred.labels = c("intercept", "VoC", "trial", "age", "VoC x trial", "VoC x age", "trial x age", "VoC x trial x age"),
          transform = NULL,
          show.est = T, 
          show.se = T, 
          show.stat = F,
          show.ci = F,
          show.re.var = F,
          show.icc = F,
          show.ngroups = F,
          show.obs = F,
          show.r2 = F,
          string.se = "SE",
          emph.p = F,
          string.pred = "Predictor",
          title = "",
          dv.labels = "")
```

## Model: Agency decisions when VoC = 0
```{r voc 0 model}

#filter data
banditTask.vocZeroTrials <- banditTask %>%
    filter(voc == 0)

#scale age
banditTask.vocZeroTrials$zAge <- scale(banditTask.vocZeroTrials$age)

# predict agency choice from utility of control, trial, linear age
agency_vocZero_byAge.mixed = mixed(agency ~  zAge + (1|subID), 
                        data = banditTask.vocZeroTrials, 
                        family = binomial, 
                        method = "LRT", control=glmerControl(optimizer="bobyqa", optCtrl=list(maxfun=1e6))) 

#display stats
agency_vocZero_byAge.mixed
summary(agency_vocZero_byAge.mixed)
```


## Plot: Sensitivity to the value of choice
```{r voc plot, fig.height = 4, fig.width = 7, unit = "in"}

VoC_plot_sub_means <- banditTask %>% 
    mutate(taskHalf = case_when(trial < 158 ~ "First Half of Task",
                                trial > 157 ~ "Second Half of Task")) %>%
    group_by(age_group, taskHalf, voc, subID) %>%
    summarize(meanSubAgency = mean(agency, na.rm = T))

VoC_plot_means <- VoC_plot_sub_means %>% 
    group_by(age_group, taskHalf, voc) %>%
    summarize(meanAgency = mean(meanSubAgency, na.rm = T),
              seAgency = sd(meanSubAgency / sqrt(n())))

#plot
VoC_plot <- ggplot(VoC_plot_means, aes(x = voc, y = meanAgency, color = age_group)) +
    facet_wrap(~taskHalf) +
    geom_point(aes(color = age_group)) + 
    geom_errorbar(aes(color = age_group, ymin = meanAgency - seAgency, ymax = meanAgency + seAgency), width = .1) + 
    geom_line(aes(group = age_group)) +
    scale_color_manual(values=c("#702963", "#c00000", "#ffae42"), name = "Age Group") +
    xlab("Value of Choice (VoC)") +
    ylab("Proportion Agency Choices") +
    geom_hline(yintercept = .5, linetype = "dashed") +
    geom_vline(xintercept = 0, linetype = "dashed") +
    voc_theme()
VoC_plot
```

## Plot: Sensitivity to value of control with continuous age 
```{r voc plot continuous age, fig.height = 3.9, fig.width = 3, unit = "in"}

#run model without age to get random effects for each participant
agency_byVOCTrial.glmer = mixed(agency ~ zVoC * zTrialOfCond + (zVoC * zTrialOfCond | subID), 
                        data = banditTask, 
                        family = binomial, 
                        method = "LRT", control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=1e6)),
                        return = "merMod") 

#get fixed effect of zVoC
VoC_fixedeff <- as.data.frame(coef(summary(agency_byVOCTrial.glmer)))$Estimate[2]
VoC_int_fixedeff <- as.data.frame(coef(summary(agency_byVOCTrial.glmer)))$Estimate[4]

#get random effects
VoC_effects <- ranef(agency_byVOCTrial.glmer)$subID %>%
    rownames_to_column(var = "subID")

#combine with age
VoC_subEffects <- banditTask %>%
    select(subID, age) %>% 
    unique() %>%
    left_join(VoC_effects, by = c("subID")) %>%
    mutate(zVoCFull = zVoC + VoC_fixedeff, 
           intFull = `zVoC:zTrialOfCond` + VoC_int_fixedeff)

#plot age by VoC effect
VoC_plot_continuousAge <- ggplot(VoC_subEffects, aes(x = age, y = zVoCFull)) +
    geom_point(color = "black") + 
    geom_smooth(method = "lm", color = "black", fill = "black") +
    voc_theme() + 
    xlab("Age") +
    ylab("VoC Effect") 
VoC_plot_continuousAge

#plot age by VoC x trial effect
VoC_plot_continuousAgeTrial <- ggplot(VoC_subEffects, aes(x = age, y = intFull)) +
    geom_point(color = "black") + 
    geom_smooth(method = "lm", color = "black", fill = "black") +
    voc_theme() + 
    xlab("Age") +
    ylab("VoC x Trial Effect") 
VoC_plot_continuousAgeTrial
```


## Summary stats: Sensitivity to value of control
```{r voc summary stats}

# What proportion of trials did participants choose agency when VoC was 0?
VoC_zero_means_sub <- banditTask %>% 
    filter(voc == 0) %>%
    group_by(subID, age_group) %>%
    summarize(meanSubAgency = mean(agency, na.rm = T))

VoC_zero_means <- VoC_zero_means_sub %>%
    summarize(meanAgency = mean(meanSubAgency, na.rm = T),
              seAgency = sd(meanSubAgency / sqrt(n())))
VoC_zero_means
```


## Agency decision reaction times
```{r agency RT data processing}

#how many RTs faster than 100 ms?
fastRTs <- banditTask %>%
    filter(agencyRT < .1) %>%
    nrow()

#486 out of 28980 trials (1.7%) of agency RTs are fast - exclude

slowRTs <- banditTask %>%
    filter(agencyRT > 30) %>%
    nrow()

#14 out of 28980 trials (< 1%) of agency RTs are very slow - exclude

banditTask.RT.data <- banditTask %>%
    filter(agencyRT > .1) %>%
    filter(agencyRT < 30) %>%
    mutate(agencyLogRT = log(agencyRT))
```

## Model: Agency RTs by VoC
```{r voc RT model}

#compute abs(voc) variable
banditTask.RT.data$absVoC <- abs(banditTask.RT.data$voc)

#scale variables
banditTask.RT.data$zAge <- scale(banditTask.RT.data$age)
banditTask.RT.data$zVoC <- scale(banditTask.RT.data$voc)
banditTask.RT.data$zAbsVoC <- scale(banditTask.RT.data$absVoC)
banditTask.RT.data$zTrialOfCond <- scale(banditTask.RT.data$trialOfCond)

agency.RT.model <- mixed(agencyLogRT ~ zAge * zAbsVoC * zTrialOfCond + (zAbsVoC * zTrialOfCond | subID),
                         data = banditTask.RT.data,
                         method = "S", 
                         control=lmerControl(optimizer="bobyqa", optCtrl=list(maxfun=1e6)))

agency.RT.model

#effects of
# age
# VoC
# trial of condition

#no interactions
```

```{r agency RTs print model stats}
agency.RT.lmer <- mixed(agencyLogRT ~ zAge * zAbsVoC * zTrialOfCond + (zAbsVoC * zTrialOfCond | subID),
                         data = banditTask.RT.data,
                         method = "S", 
                         control=lmerControl(optimizer="bobyqa", optCtrl=list(maxfun=1e6)),
                        return = "merMod")

tab_model(agency.RT.lmer, 
          pred.labels = c("intercept", "age", "VoC magnitude", "trial", "age x VoC magnitude", "age x trial", "VoC magnitude x trial", "age x VoC magnitude x trial"),
          transform = NULL,
          show.est = T, 
          show.se = T, 
          show.stat = F,
          show.ci = F,
          show.re.var = F,
          show.icc = F,
          show.ngroups = F,
          show.obs = F,
          show.r2 = F,
          string.se = "SE",
          emph.p = F,
          string.pred = "Predictor",
          title = "",
          dv.labels = "")
```


## Plot: Agency RTs by VoC
```{r agency RT by VoC, fig.height = 4, fig.width = 8, units = "in"}

agencyRT.means.sub <- banditTask.RT.data %>%
    group_by(voc, age_group, subID) %>%
    summarize(meanSubRT = mean(agencyRT),
              meanSubLogRT = mean(agencyLogRT))

#compute means and SDs
agencyRT.means <- agencyRT.means.sub  %>%
    group_by(voc, age_group) %>%
    summarize(meanRT = mean(meanSubRT),
              meanLogRT = mean(meanSubLogRT),
              seRT = sd(meanSubRT)/ sqrt(n()),
              seLogRT = sd(meanSubLogRT)/ sqrt(n()))

agencyRT.VoC.plot <- ggplot(agencyRT.means, 
                            aes(x = voc, y = meanRT)) +
    facet_wrap(~age_group) +
    geom_point(size = 3, aes(color= age_group)) +
    geom_errorbar(aes(ymin = meanRT - seRT, ymax = meanRT + seRT, color = age_group), width = .1, position = position_dodge(width = .9)) +
    scale_color_manual(values = c(color1, color2, color3)) +
    xlab("Value of Choice") +
     ylab("Mean Agency Decision Time (s)") +
    voc_theme() +
    theme(legend.position = "none")
agencyRT.VoC.plot
```

# Choice preference task 
## Choice preference task: summary stats
```{r reward sense summary stats}

# Read in data
rewardSense <- read_csv('data/processed/reward_sensitivity_task.csv') 

#combine with age
rewardSense <- full_join(rewardSense, sub_info, by = c("subID"))

# summary stats for accuracy
overallAcc <- rewardSense %>% 
    group_by(subID) %>% 
    filter(accuracy!=0) %>% 
    summarize(m=mean(correct, na.rm=T)) %>% 
    ungroup() %>% 
    summarize(meanAccuracy = mean(m), stdev = sd(m))
overallAcc

# mean = 76.9%
# stdev = 15.3%
```

## Model: Choice preference task accuracy
```{r bandit choices across by age in post-task assessment}

# first, filter data and rescale variables
rewardSense.filtered <- rewardSense %>%  
    filter(accuracy!=0)

# rescale variables of age and the true probability differences between two displayed bandits 
rewardSense.filtered$zAge <- scale(rewardSense.filtered$age)
rewardSense.filtered$zDiff<- scale(rewardSense.filtered$diff)

# run model
rewardSense.mixed <- mixed(correct~zDiff*zAge + (zDiff|subID), 
                           data= rewardSense.filtered,
                           family = binomial,
                           method = "LRT")
rewardSense.mixed 
```


# Explicit reward knowledge task 
## Explicit reward knowledge task: summary stats
```{r explicit knowledge task}

# Read in data
explicitKnow <- read_csv('data/processed/explicit_knowledge_task.csv') 

#combine with age
explicitKnow <- full_join(explicitKnow, sub_info, by = c("subID"))

explicitKnow %>% 
  group_by(subID, age) %>% 
  summarize(m = mean(error)) %>% 
  ungroup() %>% 
  summarize(meanErr = mean(m, na.rm=T), sd = sd(m,na.rm = T))
```

## Model: Explicit reward knowledge by age and true probabilities
```{r explicit knowledge model}
# predict trial-level error from true probability and age

#re-scale age and zTrueProb
explicitKnow.filtered <- explicitKnow %>%
    select(subID, age, trueProb, response, error) %>%
    drop_na()

explicitKnow.filtered$zAge <- scale(explicitKnow.filtered$age)
explicitKnow.filtered$zTrueProb <- scale(explicitKnow.filtered$trueProb)

# run model
explicitKnow_errorbyTrueProbAge.mixed <- mixed(error ~ zTrueProb*zAge + (1|subID), 
                                               data = explicitKnow.filtered,
                                               method = "S") 
explicitKnow_errorbyTrueProbAge.mixed
```

## Plot: Explicit reward knowledge
```{r plot explicit knowledge}
# plot response by bandit
explicitKnow %>% 
    ggplot(., aes(x=factor(trueProb), y=response, fill=age_group)) +
    geom_boxplot() +
    scale_fill_manual(values = c(color1, color2, color3), name = "Age Group") +
    ylab("Reported Reward Probability") +
    xlab("True Reward Probability") +
    scale_x_discrete(labels = c("10%", "30%", "50%", "70%", "90%")) +
    scale_y_continuous(breaks = c(1, 2, 3, 4, 5, 6, 7, 8, 9), 
                     labels = c("10%", "20%", "30%", "40%", "50%", "60%", "70%", "80%", "90%")) +
    voc_theme()
```
