---
title: "VoC analyses"
date: 10/1/23
output:
    html_document:
        df_print: 'paged'
        toc: true
        toc_float:
            collapsed: false
            smooth_scroll: true
        number_sections: false
        code_download: true
        self_contained: true
---
```{r chunk settings, include = FALSE}
# set chunk settings
knitr::opts_chunk$set(echo = FALSE, 
                      cache = TRUE,
                      message = FALSE,
                      warning = FALSE)
knitr::opts_chunk$set(dpi=600)
```
                      
```{r setup, include=FALSE}

# list all packages required for the analysis
list_of_packages <- c("here", "tidyverse", "afex")

# load all packages 
lapply(list_of_packages, library, character.only = TRUE)

# add theme for plotting
perri_theme <- function () {
  theme(
    panel.border = element_rect(fill = "transparent", color="gray75"),
    panel.background  = element_blank(),
    plot.background = element_blank(), 
    legend.background = element_rect(fill="transparent", colour=NA),
    legend.key = element_rect(fill="transparent", colour=NA),
    line = element_blank(),
    axis.ticks = element_line(color="gray75"),
    text=element_text(family="Avenir"),
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 15),
    title = element_text(size = 15),
    strip.background = element_blank(),
    strip.text = element_text(size=12)
  )
}

color1 = "#00b4d8"
color2 = "#0077b6"
color3 = "#03045e"

# load data
load("data/vocDataTidy.RData")
```

# Participant demographics
```{r participants plot}

# plot histogram of male and female participants
banditTask %>% mutate(wholeAge = floor(age)) %>% 
    group_by(subID, gender,wholeAge) %>% 
    distinct(subID) %>% 
  ggplot(., aes(x=wholeAge, fill=gender)) +
  geom_histogram(binwidth = 1, color="white") +
  scale_fill_manual(name="Sex",
                    labels=c("Female", "Male"),
                    values=c("#e08283", "#00b5cc")) +
  scale_y_continuous(breaks = c(2,4,6,8,10),
                   labels = c("2","4","6","8","10"),
                   limits = c(0,10)) +
  perri_theme() +
  xlab("Age") +
  ylab("Count")
```


# Agency task: Machine selection
## Model: Optimal machine choices across trials by condition and age
```{r machine choices across trials by age}

# Filter data to have only trials where people choose agency and exclude trials with 50-50 condition 
banditTask.filtered <- banditTask %>% 
    filter(agency==1, condition!="bandits5050")

# Scale continuous variables
banditTask.filtered$zAge <- scale(banditTask.filtered$age)
banditTask.filtered$zTrialOfCond <- scale(banditTask.filtered$trialOfCond)

# Mixed-effects logistic regression model
correct_byConditionTrialAge.mixed <- mixed(correct ~ condition*zTrialOfCond*(zAge) + (condition*zTrialOfCond|subID), 
                data = banditTask.filtered,
                family = binomial, 
                method = "LRT",
                control=glmerControl(optimizer="bobyqa", optCtrl=list(maxfun=1e6)))

#display model stats
correct_byConditionTrialAge.mixed 
summary(correct_byConditionTrialAge.mixed)
```

## Plot: Proportion optimal machine selections across age groups and trials
```{r plot bandit choices across trials, width = 7, height = 4, unit = "in"}

banditTaskSubMeans <- banditTask %>%
    filter(agency==1, condition!="bandits5050") %>% 
    group_by(condition, block, ageGrp, subID) %>% 
    summarize(pctCorrect = mean(correct))

banditTaskMeans <- banditTaskSubMeans %>%
    group_by(condition, block, ageGrp) %>% 
    summarize(pctCorr = mean(pctCorrect),
              se = sd(pctCorrect)/sqrt(n()))

machineSelectionPlot <- ggplot(banditTaskMeans, aes(x=block, y=pctCorr, color=condition)) +
    facet_wrap(~ageGrp) +
    geom_point(size = 3) +
    geom_jitter(data = banditTaskSubMeans,  aes(x=block, y=pctCorrect, color=condition), size = .5) +
    geom_smooth(method = "lm", aes(fill = condition)) +
    geom_hline(yintercept = .5, linetype="dashed") +
    ylab("Proportion Optimal Machine Selections") +
    xlab("Block") +
    scale_x_continuous(breaks = c(4, 8, 12)) +
    scale_fill_manual(name="Condition",
                      labels=c("70/30",
                               "90/10"),
                      values=c("#0077b6", "#03045e"), 
                      guide = guide_legend(reverse=TRUE)) +
    scale_color_manual(name="Condition",
                      labels=c("70/30",
                               "90/10"),
                      values=c("#0077b6", "#03045e"),
                     guide = guide_legend(reverse=TRUE)) +
    perri_theme() +
    theme(strip.text = element_text(size=12))
machineSelectionPlot
```


# Agency task: Agency decisions 
## Model: Agency decisions by VoC
```{r voc model}
#scale uoc
banditTask$zVoC<- scale(banditTask$uoc)

# predict agency choice from utility of control, trial, linear age
agency_byVOCTrialAge.mixed = mixed(agency ~ zVoC * zTrialOfCond * zAge + (zVoC * zTrialOfCond|subID), 
                        data = banditTask, 
                        family = binomial, 
                        method = "LRT", control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=1e6))) 

#display stats
agency_byVOCTrialAge.mixed
summary(agency_byVOCTrialAge.mixed)
```

## Model: Agency decisions when VoC = 0
```{r voc 0 model}

#filter data
banditTask.vocZeroTrials <- banditTask %>%
    filter(uoc == 0)

#scale age
banditTask.vocZeroTrials$zAge <- scale(banditTask.vocZeroTrials$age)

# predict agency choice from utility of control, trial, linear age
agency_vocZero_byAge.mixed = mixed(agency ~  zAge + (1|subID), 
                        data = banditTask.vocZeroTrials, 
                        family = binomial, 
                        method = "LRT", control=glmerControl(optimizer="bobyqa", optCtrl=list(maxfun=1e6))) 

#display stats
agency_vocZero_byAge.mixed
summary(agency_vocZero_byAge.mixed)
```

## Plot: Sensitivity to value of control
```{r VoC plot, fig.height = 4, fig.width = 7, unit = "in"}

VoC_plot_sub_means <- banditTask %>% 
    mutate(taskHalf = case_when(trial < 158 ~ "First Half of Task",
                                trial > 157 ~ "Second Half of Task")) %>%
    group_by(ageGrp, taskHalf, uoc, subID) %>%
    summarize(meanSubAgency = mean(agency, na.rm = T))

VoC_plot_means <- VoC_plot_sub_means %>% 
    group_by(ageGrp, taskHalf, uoc) %>%
    summarize(meanAgency = mean(meanSubAgency, na.rm = T),
              seAgency = sd(meanSubAgency / sqrt(n())))

#plot
VoC_plot <- ggplot(VoC_plot_means, aes(x = uoc, y = meanAgency, color = ageGrp)) +
    facet_wrap(~taskHalf) +
    geom_point(aes(color = ageGrp)) + 
    geom_errorbar(aes(color = ageGrp, ymin = meanAgency - seAgency, ymax = meanAgency + seAgency), width = .1) + 
    geom_line(aes(group = ageGrp)) +
    perri_theme() + 
    scale_color_manual(values=c("#84347C", "#B40424", "#EB6D1E"), name = "Age Group") +
    xlab("Value of Choice (VoC)") +
    ylab("Proportion Agency Choices") +
    geom_hline(yintercept = .5, linetype = "dashed") +
    geom_vline(xintercept = 0, linetype = "dashed")
VoC_plot
```

## Plot: Sensitivity to value of control with continuous age 
```{r VoC plot continuous age, fig.height = 3.9, fig.width = 3, unit = "in"}

#run model without age to get random effects for each participant
agency_byVOCTrial.glmer = mixed(agency ~ zVoC * zTrialOfCond + (zVoC * zTrialOfCond|subID), 
                        data = banditTask, 
                        family = binomial, 
                        method = "LRT", control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=1e6)),
                        return = "merMod") 

#get fixed effect of zVoC
VoC_fixedeff <- as.data.frame(coef(summary(agency_byVOCTrial.glmer)))$Estimate[2]
VoC_int_fixedeff <- as.data.frame(coef(summary(agency_byVOCTrial.glmer)))$Estimate[4]

#get random effects
VoC_effects <- ranef(agency_byVOCTrial.glmer)$subID %>%
    rownames_to_column(var = "subID")

#combine with age
VoC_subEffects <- banditTask %>%
    select(subID, age) %>% 
    unique() %>%
    left_join(VoC_effects, by = c("subID")) %>%
    mutate(zVoCFull = zVoC + VoC_fixedeff, 
           intFull = `zVoC:zTrialOfCond` + VoC_int_fixedeff)

#plot age by VoC effect
VoC_plot_continuousAge <- ggplot(VoC_subEffects, aes(x = age, y = zVoCFull)) +
    geom_point(color = "#EB6D1E") + 
    geom_smooth(method = "lm", color = "#84347C", fill = "#84347C") +
    perri_theme() + 
    xlab("Age") +
    ylab("VoC Effect") 
VoC_plot_continuousAge

#plot age by VoC x trial effect
VoC_plot_continuousAgeTrial <- ggplot(VoC_subEffects, aes(x = age, y = intFull)) +
    geom_point(color = "#EB6D1E") + 
    geom_smooth(method = "lm", color = "#84347C", fill = "#84347C") +
    perri_theme() + 
    xlab("Age") +
    ylab("VoC x Trial Effect") 
VoC_plot_continuousAgeTrial
```


## Summary stats: Sensitivity to value of control
```{r VoC summary stats}

# What proportion of trials did participants choose agency when VoC was 0?
VoC_zero_means_sub <- banditTask %>% 
    filter(uoc == 0) %>%
    group_by(subID) %>%
    summarize(meanSubAgency = mean(agency, na.rm = T))

VoC_zero_means <- VoC_zero_means_sub %>%
    summarize(meanAgency = mean(meanSubAgency, na.rm = T),
              seAgency = sd(meanSubAgency / sqrt(n())))
VoC_zero_means
```

## Plot: Agency RTs
```{r agency RT plot}

banditTask.RT.plot <- ggplot(banditTask, aes(x = agencyRT, fill = ageGrp)) + facet_wrap(~ageGrp) +
    geom_histogram(color = "black") +
    scale_fill_manual(values = c(color1, color2, color3), name = "Age Group") + 
    ylab("Number of Responses") +
    xlab("Agency Response Time (s)") +
    perri_theme()
banditTask.RT.plot

```

## Plot: Agency Log RTs
```{r agency log RT plot}

banditTask.RT.plot <- ggplot(banditTask, aes(x = log(agencyRT), fill = ageGrp)) + facet_wrap(~ageGrp) +
    geom_histogram(color = "black") +
    scale_fill_manual(values = c(color1, color2, color3), name = "Age Group") + 
    ylab("Number of Responses") +
    xlab("Log Agency Response Time (s)") +
    perri_theme()
banditTask.RT.plot

```

```{r agency RT data processing}

#how many RTs faster than 100 ms?
fastRTs <- banditTask %>%
    filter(agencyRT < .1) %>%
    nrow()

#486 out of 28980 trials (1.7%) of agency RTs are fast - exclude

slowRTs <- banditTask %>%
    filter(agencyRT > 30) %>%
    nrow()

#14 out of 28980 trials (< 1%) of agency RTs are very slow - exclude

banditTask.RT.data <- banditTask %>%
    filter(agencyRT > .1) %>%
    filter(agencyRT < 30) %>%
    mutate(agencyLogRT = log(agencyRT))
```

## Model: Agency RTs by VoC
```{r agency RT model}

#scale variables
banditTask.RT.data$zAge <- scale(banditTask.RT.data$age)
banditTask.RT.data$absVoC <- abs(banditTask.RT.data$uoc)
banditTask.RT.data$zAbsVoC <- scale(banditTask.RT.data$absVoC)
banditTask.RT.data$zVoC <- scale(banditTask.RT.data$uoc)
banditTask.RT.data$zTrialOfCond <- scale(banditTask.RT.data$trialOfCond)

agency.RT.model <- mixed(agencyLogRT ~ zAge * zAbsVoC * zTrialOfCond + (zAbsVoC * zTrialOfCond | subID),
                         data = banditTask.RT.data,
                         method = "S", 
                         control=lmerControl(optimizer="bobyqa", optCtrl=list(maxfun=1e6)))

agency.RT.model
summary(agency.RT.model)

#effects of
# age
# VoC
# trial of condition

#no interactions
```

## Plot: Agency RTs by VoC
```{r agency RT by VoC, fig.height = 4, fig.width = 4, units = "in"}

agencyRT.means.sub <- banditTask.RT.data %>%
    group_by(uoc, ageGrp, subID) %>%
    summarize(meanSubRT = mean(agencyRT),
              meanSubLogRT = mean(agencyLogRT))

#compute means and SDs
agencyRT.means <- agencyRT.means.sub  %>%
    group_by(uoc) %>%
    summarize(meanRT = mean(meanSubRT),
              meanLogRT = mean(meanSubLogRT),
              seRT = sd(meanSubRT)/ sqrt(n()),
              seLogRT = sd(meanSubLogRT)/ sqrt(n()))

agencyRT.VoC.plot <- ggplot(agencyRT.means, 
                            aes(x = uoc, y = meanRT)) +
    geom_point(size = 3, color = color1) +
    geom_errorbar(aes(ymin = meanRT - seRT, ymax = meanRT + seRT), width = .1, position = position_dodge(width = .9), color = color1) +
    xlab("Value of Choice") +
     ylab("Mean Agency Decision Time (s)") +
    perri_theme()
agencyRT.VoC.plot
```


# Choice preference task 
## Choice preference task: summary stats
```{r reward sense summary stats}
# summary stats for accuracy
overallAcc <- rewardSense %>% 
    group_by(subID) %>% 
    filter(accuracy!=0) %>% 
    summarize(m=mean(correct, na.rm=T)) %>% 
    ungroup() %>% 
    summarize(meanAccuracy = mean(m), stdev = sd(m))
overallAcc

# mean = 76.9%
# stdev = 15.3%
```

## Model: Choice preference task accuracy
```{r bandit choices across by age in post-task assessment}

# first, filter data and rescale variables
rewardSense.filtered <- rewardSense %>%  filter(accuracy!=0)

# rescale variables of age and the true probability differences between two displayed bandits 
rewardSense.filtered$zAge <- scale(rewardSense.filtered$age)
rewardSense.filtered$zDiff<- scale(rewardSense.filtered$diff)

# run model
rewardSense.mixed <- mixed(correct~zDiff*zAge + (zDiff|subID), 
                           data= rewardSense.filtered,
                           family = binomial,
                           method = "LRT")
rewardSense.mixed 
```


# Explicit reward knowledge task 
## Explicit reward knowledge task: summary stats
```{r}
# summary stats for error
# mean = 1.43
# stdev = .872

explicitKnow %>% 
  group_by(subID, age) %>% 
  summarize(m = mean(error)) %>% 
  ungroup() %>% 
  summarize(meanErr = mean(m, na.rm=T), sd = sd(m,na.rm = T))
```

## Model: Explicit reward knowledge by age and true probabilities
```{r explicit knowledge model}
# predict trial-level error from true probability and age

#re-scale age and zTrueProb
explicitKnow.filtered <- explicitKnow %>%
    select(subID, age, trueProb, response, error) %>%
    drop_na()

explicitKnow.filtered$zAge <- scale(explicitKnow.filtered$age)
explicitKnow.filtered$zTrueProb <- scale(explicitKnow.filtered$trueProb)

# run model
explicitKnow_errorbyTrueProbAge.mixed <- mixed(error ~ zTrueProb*zAge + (1|subID), 
                                               data = explicitKnow.filtered,
                                               method = "S") 
explicitKnow_errorbyTrueProbAge.mixed
```

## Plot: Explicit reward knowledge
```{r plot explicit knowledge}
# plot response by bandit
explicitKnow %>% 
    ggplot(., aes(x=bandit,y=response, fill=ageGrp)) +
    geom_boxplot() +
    scale_fill_manual(values = c(color1, color2, color3), name = "Age Group") +
    ylab("Reported Reward Probability") +
    xlab("True Reward Probability") +
    scale_x_discrete(labels = c("10%", "30%", "50%", "50%", "70%", "90%")) +
    scale_y_continuous(breaks = c(1, 2, 3, 4, 5, 6, 7, 8, 9), 
                     labels = c("10%", "20%", "30%", "40%", "50%", "60%", "70%", "80%", "90%")) +
    perri_theme()
```
